---
title: "การสอนโปรแกรม R"
author: "เฉลิมพงษ์ คงเจริญ"
output:
  word_document: default
  html_document: default
  pdf_document: default
---


# Why R?
-	Powerful and flexible
-	Free(open source)
-	Extensive add-on software(package)
-	Designed for statistical computing
-	High level language

# Installing R
-	Install the latest version from: http//cran.r-project.org
-	Install R Studio (Makes R easier) from http://www.rstudio.com

นอกจากนี้ เราสามารถใช้ Rstudio บน internet โดยไม่ต้องลงโปรแกรมผ่าน https://rstudio.cloud/

# Working with R
- The R Console ‘interprets’ whatever you type
    + Calculator
    + Creating variables
    + Applying function
-	“Analysis” Script+ Interactive Exploration
    + Static copy of what you did (reproducibility)
    + Try things out interactively, then add to your script
-	R revolves around functions
    + Commands that take input, performs computations, and returns results
    + Many come with R Base (installation), but people write external functions you can download and use

## R as a calculator
```{r}
2+2
2*4
2^3
2+(2*3)^2
```

##R variables
-	You can create variables from within the R environment and from files on your computer
-	R uses “=” or “<-“ to assign values to a variable name
-	Variable names are case-sensitive, i.e. X and x are different.
```{r}
a<-2*3
2*a
```

The ‘combine’ function – the function c( ) collects/combines/joins single R objects into a vector of R objects.

```{r}
b<- c(2,3,4)
b
```

#What is a function?
The c( ) command is called a function: it takes inputs and gives an output.
-	In R, functions always go function(input) or name of function, then parentheses.  The input can be many different things, such as function(x,y,z)

```{r}
x<-c(5,9,13,4)
mean(x)
```

```{r}
y<-rnorm(10)
y
set.seed(123)
y<-rnorm(10)
y
```
#Graphic

```{r}
x<-rnorm(100)
plot(x)
```

#R Help
If you know the name of a function, help(function name)

#Data Classes:
	One dimensional classes(‘vectors’)
-	Character
-	Numeric
-	Integer
-	Factor
-	Logical
Two dimensional classes
-	data.frame: traditional ‘Excel’ spreadsheet
-	Matrix: two-dimensional data, composed of rows and columns

##Vector functions
```{r}
z<-1:100
head(z)
tail(z)
str(z)
```


##Data Subsetting
```{r}
x1<- 10:20
length(x1)
x1[1]
x1[3:4]
x1[c(1,5,7)]
```

##Matrices
```{r}
m<-1:9
m
mat<-matrix(m,nrow=3)
mat
nrow(mat)
ncol(mat)
dim(mat)
```

##Data Selection
Matrices have tow ‘slots’ you can use to select data, which represent rows and columns, that are separated by a comma, so the syntax is matrix[row, column]

```{r}
mat[1,1]
mat[1, ]
mat[ ,1]
```

##Data Frame

```{r}
t<-data.frame(x=c(1,2,3), y=c(10,20,30), z=c(100,200,300))
t
mean(t$y)
mean(t[["z"]])
```

##List

```{r}
L<-list(one=1, two=c(1,2), five=seq(1,4, length=5))
L
```
##Graphic 

```{r}
set.seed(123)
y<-rnorm(100)
plot(y)
plot(y, type="l", col="blue", ylab="value", xlab="observations", main="Generated Data of N(0,1)")
```
```{r}
t<-data.frame(x=c(1,2,3), y=c(4,5,6), z=c(7,8,9))
plot(t$x, type="l",ylim=c(0,10), ylab="value")
lines(t$y, col="red")
points(t$z, pch=20, cex=2, col="blue")
```


```{r}
ptt <- read.csv("https://raw.githubusercontent.com/chaleampong/EC435/master/ptt_d_02_19.csv", stringsAsFactors=FALSE)
class(ptt)
names(ptt)
str(ptt)
head(ptt$Price)
```

#More on Packages
Packages are add-ons that are commonly written by users comprised of functions, data and vignettes.
-	Use library( ) or require( ) to load the package into memory so you can use its functions.
-	Install packages using install.packages(“PackageName”) .
-	Use help(package=”PackageName”) to see what contents the package has.


#Import Data
In RStudio, we can import data files CSV, Excel, SPSS, SAS, Stata

#Saving R data
It’s very useful to be able to save collections of R objects for future analyses. Save(…, file=”name.rda”)
When you close R, you might notice the prompt about saving your workspace in “.Rdata”.

##Saving and Importing Data from text file
```{r}
#Saving Data
d<-data.frame(a=c(3,4,5), b=c(12,13,14))
d
write.table(d, file="ts0.txt", row.names=FALSE)
#Importing Data
d2<-read.table(file="ts0.txt", header=TRUE)
d2
```

#การคำนวณผลได้ตอบแทน

```{r}
ptt<-read.csv("G:/My Drive/teaching/ec435/book_project/data/ptt_d_02_19.csv")
head(ptt)
ptt.lret<-diff(log(ptt$Price))
n<-length(ptt$Price)
ptt.sret<-(ptt$Price[2:n]-ptt$Price[1:n-1])/(ptt$Price[1:n-1])
ptt$sret<-c(NA,ptt.sret)
head(ptt)
write.csv(ptt,file="ptt_ret.csv", row.names=FALSE)
```

##Adding to data frames
```{r}
ptt$lret<-c(NA,ptt.lret)
head(ptt)
```

#Adding rows and columns

```{r} 
m1<-matrix(1:9, nrow=3)
m2<-matrix(10:18, nrow=3)
cbind(m1,m2)
rbind(m1,m2)
```

#Other manipulations
-	abs(x)
-	sqrt(x)
-	log(x)
-	log10(x)
-	exp(x)

#Data Summarization
##Basic statistical summarization
- mean(x)
- sd(x)
- median(x)
- quantile(x)
- range(x)

##Basic summarization plots
- plot(x,y)
- hist(x)
- plot(density(x))

#Basic plot
```{r}
plot(ptt$Price, type="l", main="Daily price of PTT", xlab="Date", 
ylab="Price")
hist(ptt$lret)
plot(density(ptt$lret, na.rm=TRUE))
```

#Probability Distributions
-	r for random number generation [e.g. rnorm( )]
-	d for density [e.g. dnorm( )]
-	p for probability [e.g. pnorm( )]
-	q for quantile [e.g. qnorm( )]


#Basic Statistics and tests

```{r}
mean(ptt$lret)
mean(ptt$lret, na.rm = TRUE)
sd(ptt$lret, na.rm= TRUE)
library(fBasics)
s3<-skewness(ptt$lret, na.rm= TRUE)
s3
T<-nrow(ptt)
t3<-s3/sqrt(6/T)
t3
k4<-kurtosis(ptt$lret, na.rm= TRUE)
k4
t4<-k4/sqrt(24/T)
t4
normalTest(ptt$lret,method=c("jb"))

basicStats(ptt$lret)
```
