---
title: "การสอนโปรแกรม R"
author: "เฉลิมพงษ์ คงเจริญ"
output: word_document
      
---


#Why R?
-	Powerful and flexible
-	Free(open source)
-	Extensive add-on software(package)
-	Designed for statistical computing
-	High level language

#Installing R
-	Install the latest version from: http//cran.r-project.org
-	Install R Studio (Makes R easier) from http://www.rstudio.com

นอกจากนี้ เราสามารถใช้ Rstudio บน internet โดยไม่ต้องลงโปรแกรมผ่าน https://rstudio.cloud/

#Working with R
- The R Console ‘interprets’ whatever you type
    + Calculator
    + Creating variables
    + Applying function
-	“Analysis” Script+ Interactive Exploration
    + Static copy of what you did (reproducibility)
    + Try things out interactively, then add to your script
-	R revolves around functions
    + Commands that take input, performs computations, and returns results
    + Many come with R Base (installation), but people write external functions you can download and use

##R as a calculator
```{r}
2+2
2*4
2^3
2+(2*3)^2
```

##R variables
-	You can create variables from within the R environment and from files on your computer
-	R uses “=” or “<-“ to assign values to a variable name
-	Variable names are case-sensitive, i.e. X and x are different.
```{r}
x<-2
x
```

The ‘combine’ function – the function c( ) collects/combines/joins single R objects into a vector of R objects.

```{r}
x<- c(1,4,6,8)
x
```

#What is a function?
The c( ) command is called a function: it takes inputs and gives an output.
-	In R, functions always go function(input) or name of function, then parentheses.  The input can be many different things, such as function(x,y,z)

#R Help
If you know the name of a function, help(function name)

#Data Classes:
	One dimensional classes(‘vectors’)
-	Character
-	Numeric
-	Integer
-	Factor
-	Logical
Two dimensional classes
-	data.frame: traditional ‘Excel’ spreadsheet
-	Matrix: two-dimensional data, composed of rows and columns

##Vector functions
```{r}
z<-1:100
head(z)
tail(z)
str(z)
```


##Data Subsetting
```{r}
x1<- 10:20
length(x1)
x1[1]
x1[3:4]
x1[c(1,5,7)]
```

##Matrices
```{r}
m<-1:9
m
mat<-matrix(m,nrow=3)
mat
nrow(mat)
ncol(mat)
dim(mat)
```

##Data Selection
Matrices have tow ‘slots’ you can use to select data, which represent rows and columns, that are separated by a comma, so the syntax is matrix[row, column]

```{r}
mat[1,1]
mat[1, ]
mat[ ,1]
```

##Data Frame

```{r}
pttstock <- read.csv("https://raw.githubusercontent.com/chaleampong/EC435/master/pttstock.csv", stringsAsFactors=FALSE)
class(pttstock)
names(pttstock)
str(pttstock)
head(pttstock$price)
```

#More on Packages
Packages are add-ons that are commonly written by users comprised of functions, data and vignettes.
-	Use library( ) or require( ) to load the package into memory so you can use its functions.
-	Install packages using install.packages(“PackageName”) .
-	Use help(package=”PackageName”) to see what contents the package has.


#Import Data
In RStudio, we can import data files CSV, Excel, SPSS, SAS, Stata

#Saving R data
It’s very useful to be able to save collections of R objects for future analyses. Save(…, file=”name.rda”)
When you close R, you might notice the prompt about saving your workspace in “.Rdata”.

#การคำนวณผลได้ตอบแทน

```{r}
ptt.lret<-diff(log(pttstock$price))
n<-length(pttstock$price)
ptt.sret<-(pttstock$price[2:n]-pttstock$price[1:n-1])/(pttstock$price[1:n-1])
pttstock$sret<-c(NA,ptt.sret)
head(pttstock)
write.csv(pttstock,file="pttstock.csv", row.names=FALSE)
```

##Adding to data frames
```{r}
pttstock$lret<-c(NA,ptt.lret)
head(pttstock)
```

#Adding rows and columns

```{r} 
m1<-matrix(1:9, nrow=3)
m2<-matrix(10:18, nrow=3)
cbind(m1,m2)
rbind(m1,m2)
```

#Other manipulations
-	abs(x)
-	sqrt(x)
-	log(x)
-	log10(x)
-	exp(x)

#Data Summarization
##Basic statistical summarization
- mean(x)
- sd(x)
- median(x)
- quantile(x)
- range(x)

##Basic summarization plots
- plot(x,y)
- hist(x)
- plot(density(x))

#Basic plot
```{r}
plot(pttstock$price, type="l", main="Daily price of PTT", xlab="Date", 
ylab="Price")
hist(pttstock$lret)
plot(density(pttstock$lret, na.rm=TRUE))
```

#Probability Distributions
-	r for random number generation [e.g. rnorm( )]
-	d for density [e.g. dnorm( )]
-	p for probability [e.g. pnorm( )]
-	q for quantile [e.g. qnorm( )]


#Basic Statistics and tests

```{r}
mean(pttstock$lret)
mean(pttstock$lret, na.rm = TRUE)
sd(pttstock$lret, na.rm= TRUE)
library(fBasics)
s3<-skewness(pttstock$lret, na.rm= TRUE)
s3
T<-nrow(pttstock)
t3<-s3/sqrt(6/T)
t3
k4<-kurtosis(pttstock$lret, na.rm= TRUE)
k4
t4<-k4/sqrt(24/T)
t4
normalTest(pttstock$lret,method=c("jb"))

basicStats(pttstock$lret)
```

#Package zoo และ PerformanceAnalytics
```{r}
library(PerformanceAnalytics)
library(xts)
#convert data.frame to xts
pttstock <- read.csv("https://raw.githubusercontent.com/chaleampong/EC435/master/pttstock.csv", stringsAsFactors=FALSE)
class(pttstock$date)
pttstock$date<-as.Date(as.character(pttstock$date), "%m/%d/%Y")
pttstock.xts<-xts(pttstock$price, order.by=pttstock$date)
head(pttstock.xts)
plot(pttstock.xts)
pttstock.xts.sret<-Return.calculate(pttstock.xts, method=c("discrete"))
plot(pttstock.xts.sret, main = "simple return of PTT stock")
pttstock.xts.lret<-Return.calculate(pttstock.xts, method=c("log"))
plot(pttstock.xts.lret, main="log return of PTT stock")
```

การทำงานบน R สามารถเปลี่ยน Folder(Directory) ที่ทำงานได้ โดยเราสามารถตรวจสอบด้วยคำสั่ง getwd( ) และเปลี่ยนด้วยคำสั่ง setwd(“Folder”)

```{r}
getwd()
#setwd()
```

เราสามารถ Save Data.Frame ได้ด้วยคำสั่ง write.csv โดยไฟล์ดังกล่าวจะปรากฏใน Folder ที่เราทำงาน

```{r}
write.csv(pttstock, file="pttreturn.csv")
```

นอกจากนี้เราสามารถ Save ทุกอย่างที่เราทำงานได้ด้วยคลิก File>Save และตั้งชื่อไฟล์แล้ว Save (ไฟล์จะอยู่ในนามสกุล .Rdata)
```{r}
save.image(file="pttsession.Rdata")
```
